---
title: "WaPo Police Fatal Shootings"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
library(ggthemes)
library(RColorBrewer)
library(lattice)
library(tidycensus)
library(corrr)
library(janitor)

wapo_police_shootings <- read_csv("data-police-shootings-master/fatal-police-shootings-data.csv")
```

Question: Show the basic exploratory analysis to demonstrate an understanding of the dataset. 

Answer: There are various columns that give background on each fatal shooting. One of those columns include the date of the shooting. We noticed that the most recent dates have less information like their race, age, name and the fact that the body cameras weren't on at the time of the shooting. Under the "flee" column, information about that is unavailable/come up as NA for some recent shootings. 

There's a "signs_of_mental_illness" column, which also shows, through true or false, whether the victim displayed potential mental health illnesses before their death. There's also a longtitude column and a latitude column, which gives the precise location of where the victims were shot. 

A few questions we have is what NA means, especially for columns like "flee" and "armed" where they have entries like "not fleeing" and "unarmed."

Here is some code to show some basic findings...: 

```{r}
wapo_police_shootings %>% 
  group_by(flee) %>% 
  summarise(total = n()) #Majority of the victims were not fleeing.

  sum(is.na(wapo_police_shootings$name)) #269 NA names. 
  
wapo_police_shootings %>% 
  group_by(age) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total))  #NA under age comes up the most, 31 was after. 
  #arrange(desc(age))

wapo_police_shootings %>% 
  group_by(state) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total)) #California has the most shootings.

wapo_police_shootings %>% 
  group_by(city) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total)) #Los Angeles has the most shootings.

wapo_police_shootings %>% 
  group_by(armed) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total)) #Majority of the victims were armed with a gun.

wapo_police_shootings %>% 
  group_by(body_camera) %>% 
  summarise(total = n()) #7.2% of officers had their body camera on at the time of the shooting.

wapo_police_shootings %>% 
  group_by(threat_level) %>% 
  summarise(total = n()) #Majority of the victims had a threat level of "attack."

wapo_police_shootings %>% 
  group_by(gender) %>% 
  summarise(total = n()) #95% of victims were males. 

wapo_police_shootings %>% 
  group_by(race) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total)) #Majority of victims were white, followed by Black, followed by Hispanic. Also, no way to indicate more than one race. 
```

Question #1: Which state had the most fatal shootings per 100,000 population?

New Mexico and Alaska stand apart from all other states with 6.4 police shootings per 100,000 residents. Why do these two states stand apart, with the next highest rate below 5 per 100,000 people? Both could be said to have disbursed, mostly rural populations, but they're not especially different than other Western states in that regard. What do New Mexico and Alaska have in common with each other, that's different from other states, that causes them to lead the nation in police shootings? 

```{r}
# To normalize the rate of police shootings by state, we calculate the number of police shootings per 100,000 residents. 

# group shootings by state
wapo_police_shootings_bystate <- wapo_police_shootings %>%
  group_by(state) %>%
  summarise(shootings = n())

# load census variables
vars <- load_variables(year = 2019,
                       dataset = "acs1",
                       cache = TRUE)

# get state populations; join with state 2-letter abbreviations
state_pops <- get_acs(geography = "state",
                      variables = "B01001_001")

state_pops <- state_pops %>%
  select(-variable, -moe)
state_pops

state_abbreviations <- read_csv("data-police-shootings-master/name-abbr.csv")
state_abbreviations

state_pop_abbrev <- state_pops %>%
  inner_join(state_abbreviations, by = c("NAME" = "state"))

wapo_police_shootings_per100k <- state_pop_abbrev %>%
  left_join(wapo_police_shootings_bystate, by = c("abbreviation" = "state")) %>%
  mutate(shootings_percapita = shootings/estimate) %>%
  mutate(shootings_per100k = shootings_percapita * 100000) %>%
  arrange(desc(shootings_per100k))

wapo_police_shootings_per100k

```

Question #2: What type of gun-law does each state have? What about New Mexico? 

There's a stronger significance and bigger difference in means between states when open carry laws are compared. States with no restrictions on open carry (also known as permitless open carry), have an average of 2.8 shootings per 100,000 residents, meaningfully higher than the 1.8 average shootings per 100,000 people in the states with any restriction on open carry. (In this analysis, any restriction includes states with statewide or local licensing requirements, open carry only permitted in limited circumstances, or no open carry permitted.) Permitless open carry drives a higher rate of police shootings, among those incidents collected by the Washington Post. 

```{r}
# Do gun laws effect the rate of shootings by state? The analysis below looks at concealed carry and open carry laws by state (leaning on Wikipedia summaries). It divides states in to two sets -- unrestricted (concealed carry or open carry) vs any level of restriction (concealed carry or open carry) and asks whether there's a significant difference in shootings between unrestricted and restricted groupings.

# States with unrestricted (aka constitutional) concealed carry.
# https://en.wikipedia.org/wiki/Concealed_carry_in_the_United_States

unrestricted_concealed_carry <- wapo_police_shootings %>%
  filter(state %in% c("AK", "AZ", "AR", "ID", "IA", "KS", "KY", "ME", "MS", "MT", "NH", "ND", "OK", "SD", "TN", "TX", "UT", "VT", "WV","WY")) %>%
  group_by(state) %>%
  summarise(shootings = n())

unrestricted_concealed_carry

# Inverse - states with any kind of restriction on concealed carry. 
# https://en.wikipedia.org/wiki/Concealed_carry_in_the_United_States 

anyrestricted_concealed_carry <- wapo_police_shootings %>%
  filter(!state %in% c("AK", "AZ", "AR", "ID", "IA", "KS", "KY", "ME", "MS", "MT", "NH", "ND", "OK", "SD", "TN", "TX", "UT", "VT", "WV","WY")) %>%
  group_by(state) %>%
  summarise(shootings = n())

anyrestricted_concealed_carry


# calculate per 100k shootings
state_pop_shootings_unrestricted <- state_pop_abbrev %>%
  left_join(unrestricted_concealed_carry, by = c("abbreviation" = "state"))

state_pop_shootings_unrestricted <- state_pop_shootings_unrestricted %>%
  filter(shootings > 0) %>%
  mutate(shootings_percapita = shootings/estimate) %>%
  mutate(shootings_per100k = shootings_percapita * 100000) %>%
  arrange(desc(shootings_per100k))

state_pop_shootings_unrestricted 

mean(state_pop_shootings_unrestricted$shootings_per100k)

anyrestricted_concealed_carry

state_pop_shootings_anyrestrict <- state_pop_abbrev %>%
  left_join(anyrestricted_concealed_carry, by = c("abbreviation" = "state"))

state_pop_shootings_anyrestrict <- state_pop_shootings_anyrestrict %>%
  filter(shootings > 0) %>%
  mutate(shootings_percapita = shootings/estimate) %>%
  mutate(shootings_per100k = shootings_percapita * 100000) %>%
  arrange(desc(shootings_per100k))

state_pop_shootings_anyrestrict

mean(state_pop_shootings_anyrestrict$shootings_per100k)



# We find a difference in the average (mean) rate of shootings between states with unrestricted concealed carry laws and any restrictions on concealed carry. States with unrestricted laws have an average of 2.9 shootings per 100,000 people in the database, compared with 2.0 per 100,000 people for states with any kind of restriction on concealed carry.

# "Any restriction" includes states that require a license and/or permit to carry a concealed weapon, and those that do not allow carrying of concealed weapons. 

# However, it's also interesting that the two states with the highest rate of police shootings per 100,000 people, which we identified in the per capita analysis above, don't belong to the same group. Alaska, with 6.4 shootings per 100,000 people is an unrestricted open carry state. New Mexico, also with 6.4 shootings per 100,000 people, is not an unrestricted concealed carry state. 

# Is the difference between the unrestricted and restricted groups statistically significant?


# add column with boolean for restrictions on concealed carry
state_pop_shootings_unrestricted <- state_pop_shootings_unrestricted %>%
  add_column(conc_carry_restrict = 0)
  
state_pop_shootings_anyrestrict <- state_pop_shootings_anyrestrict %>%
  add_column(conc_carry_restrict = 1)

shootings_concealed_carry <- rbind(state_pop_shootings_unrestricted, state_pop_shootings_anyrestrict)

t.test(shootings_per100k ~ conc_carry_restrict, data = shootings_concealed_carry)

# The t-test suggests that higher average rate of police shootings in states with unrestricted concealed carry laws is significant (at p-value = 0.0161). 

# Open Carry

# In a similar manner, we divide the states by whether they have unrestricted open carry laws, or any restriction on open carry, to see if it has a significant relationship to the average number of shootings by state. 


# https://en.wikipedia.org/wiki/Open_carry_in_the_United_States
# add a column for open carry laws

shootings_gun_laws <- shootings_concealed_carry %>%
  add_column(open_carry_restrict = 0)

# set open_carry_restrict to 1 for states that do not have "permissive" aka permitless open carry laws
# 21 total

shootings_gun_laws <- shootings_gun_laws %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "CA", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "CO", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "CT", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "DC", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "FL", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "GA", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "HI", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "IL", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "IN", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "MD", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "MA", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "MN", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "MO", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "NE", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "NJ", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "NY", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "ND", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "OR", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "PA", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "RI", 1, open_carry_restrict)) %>%
  mutate(open_carry_restrict = ifelse(abbreviation == "SC", 1, open_carry_restrict))

t.test(shootings_per100k ~ open_carry_restrict, data = shootings_gun_laws)
```

Question #3: In what age range do the most shootings occur in New Mexico? Which age has the most shootings? Does the age have any correlation to the other states? 

```{r}
new_mex_shootings <- wapo_police_shootings %>%
  filter(state == "NM")

nm_age_shootings <- new_mex_shootings %>%
  group_by(age) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

# New Mexico Shootings by age
national_shootings <- wapo_police_shootings %>%
  group_by(age) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# National shootings by age
pct_thirtysix <- c((10/134)*100)
pct_twentyeight <- c((6/134)*100)
shootings <- c(134)
national_pct_thirtysix <- c((190/6735)*100)
national_pct_thirtyeight <- c((165/6735)*100)

all_shootings <- c (6735)

#Want to compare the cor test with New Mexico and from a national perspective
nm_age_pcts <- data.frame(pct_thirtysix, pct_twentyeight, shootings)

all_age_pct <- data.frame(national_pct_thirtysix, national_pct_thirtyeight, all_shootings)

cor.test(nm_age_pcts$pct_thirtysix, all_age_pct$national_pct_thirtysix)

#need some assistance here
```

Question #4: What about those who flee or don't flee? Which type of flee has the most amount? Compare and contrast with the other states...

```{r}
# National percentage:
wapo_police_shootings %>% 
  group_by(flee) %>% 
  summarise(total = n())

# New Mexico numbers
nm_shootings <- wapo_police_shootings %>% 
  filter(state == "NM")

nm_flee_summarized <- nm_shootings %>% 
  group_by(flee) %>% 
  summarise(total = n())

flee_pct <- c("perct_car", "perct_foot", "perct_not_flee")
new_mexico <- c((29/5951)*100, (18/5951)*100, (72/5951)*100)
national <- c((1088/5951)*100, (861/5951)*100, (4002/5951)*100)

flee_summarized <- data.frame(flee_pct, new_mexico, national)

flee_summarized %>%
  ggplot() +
  geom_point(aes(x=new_mexico,y=national)) +
  geom_smooth(aes(x=new_mexico,y=national), method="lm")

# Looking for the p value and t-test.

flee_summarized %>%
  select(-flee_pct)
  correlate() %>%
  select(term, new_mexico)

```

Question #5: What about officers with body cameras and officers who don't? Which one has the most? Is there any correlation between the body camera and being fatally shot?

```{r}
# Note: The following is similar to what I did in the previous question. So, same exact explanations, same outcomes.   

# What kinds of body camera entries are there, and which one has the most?
nm_body_cam <- nm_shootings %>%
  group_by(body_camera) %>% 
  summarise(total = n())
# True or False. True meaning the officer did have his body camera on, false meaning the officer did not have his body camera on. False has the most - 115. 

pct_body_cam <- c((115/134)*100)
pct_non_body_cam <- c((19/134)*100)
shootings_cam <- c(134)

nm_body_cam_pcts <- data.frame(pct_body_cam, pct_non_body_cam, shootings_cam)

# Attempting linear regression...
nm_body_cam_pcts %>%
  ggplot() +
  geom_point(aes(x=shootings_cam,y=pct_body_cam)) +
  geom_smooth(aes(x=shootings_cam,y=pct_body_cam), method="lm")

# Looking for the p value and t-test.
nm_body_cam_pcts %>% 
  correlate() %>% 
  select(term, shootings_cam)

```
