---
title: "lab_02"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into multiple options: under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus. How many loans are in each group?
**A1** The amount of loans under 10,000 are 9,310, between 10,000 and 50,000 are 15,372, between 50,000 and 100,000 are 3,978, between 100,000 and 500,000 are 4,152, between 500,000 and 1,000,000 are 540, and more than 1,000,000 are 356. 

```{r}
# Write the code to produce your answer here
# Like so: filter - mutate - group_by - summarise - arrange. Our question asks for one column from the data, "amount." Hence,  

wv_ppp_loan_amount_categories <- west_virginia_ppp %>%
  select(amount) %>%
  mutate(label = case_when(
    amount < 10000 ~ "under_10k", 
    amount >= 10000 & amount < 50000 ~ "btw_10k_50k",
    amount >= 50000 & amount < 100000 ~ "btw_50k_100k",
    amount >= 100000 & amount < 500000 ~ "btw_100k_500k",
    amount >= 500000 & amount < 1000000 ~ "btw_500k_1m",
    amount >= 1000000 ~ "more_than_1m")) %>% 
  group_by(label) %>% 
  summarise(total = n()) 
```

**Q2** Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** Harvest Small Business Finance, LLC has the most loans with undisbursed amounts greater than 0.

```{r}
# Write the code to produce your answer here
# Like so: filter - group_by - summarise - arrange. In regards to the question: filter - "undisbursed amounts greater than 0." Meaning filter(undisbursed_amount > 0). Group_by (what I wanna see) - "which servicing lenders" Meaning: group_by(servicing_lender_name). Summarise (does counting, total of things, also creates a new column) - counts the number of rows " Meaning: summarise (total = n()) Arrange - "the most loans," meaning - arrange(desc(amount)). 

west_virginia_ppp %>% 
  filter(undisbursed_amount > 0) %>% 
  group_by(servicing_lender_name) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total)) 
```

**Q3** How would you describe the top 10 list of those lenders? What do they have in common or not? (You may want to look up some of the top ones.)
**A3** They are all lenders/banks.

**Q4** Which month in 2021 had the most loans approved?
**A4** March

```{r}
# Write the code to produce your answer here
# Like so: filter - mutate - group_by - summarise - arrange. However, because we need to add another column and add new data - something that summarise can't do. That means mutating, "using the data you have to answer a new question" that can't be immediately answered because it isn't outright in the data. Hence,
# From the question we are looking for the "month in 2021" with the "most loans approved." Colnames we are going to use from the data are "date_approved." We are looking for the number of loans (how many there are), not the amount of loans (how much they are together). This means we will need to add columns - summarise. 
# The code: west_virginia_ppp %>% ---> the dataset we're using and the pipe.
#           mutate(Month = floor_date(date_approved, "month")) ---> "mutate" to create column and make                changes, "Month" is the new colname we will assign it to, "floor_date" is what we use                   to assign the colname, and what will help us 'round down' our dates, "date_approved" is                 the colname and information we need to make it possible, ""month"" is the unit from the                 date that we a looking for. 
#           group_by(Month) ---> "group_by" gives us the column and what we are grouping the new mutated               data with. "Month" is our new colname from the previous line of code. 
#           summarise(total = n()) ---> "summarise" counts the rows/how many loans rather than how much.               "total" is an assigned colname that gives us that summary of how many, and the column                    name. "n()" is that number exactly that is assigned to the total. 
#           arrange(desc(total)) ---> "arrange" simply arranges the data. Because we need the most,                   "desc" will arrange the data from greatest to least. "total" is the new colname that will                be arranged to give us the most loans for the month we are looking for.  

west_virginia_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>% 
  group_by(Month) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total))
```
