---
title: "lab_02"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into multiple options: under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus. How many loans are in each group?
**A1** [write your answer here]

```{r}
# Write the code to produce your answer here
# This is mutating

west_virginia_ppp %>% 
  

```

**Q2** Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** 

```{r}
# Write the code to produce your answer here
# Like so: filter - group_by - summarise - arrange. In regards to the question: filter - "undisbursed amounts greater than 0." Meaning filter(undisbursed_amount > 0). Group_by (what I wanna see) - "which servicing lenders" Meaning: group_by(servicing_lender_name). Summarise (does counting, total of things, also creates a new column) - counts the number of rows " Meaning: summarise (total = n()) Arrange - "the most loans," meaning - arrange(desc(amount)). 

west_virginia_ppp %>% 
  filter(undisbursed_amount > 0) %>% 
  group_by(servicing_lender_name) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total)) 
```

**Q3** How would you describe the top 10 list of those lenders? What do they have in common or not? (You may want to look up some of the top ones.)
**A3** [write your answer here]

**Q4** Which month in 2021 had the most loans approved?
**A4** [write your answer here]

```{r}
# Write the code to produce your answer here

```
