---
title: "lab_08"
author: "derek willis"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
install.packages("tidycensus")
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
census_api_key("dddd8ba3acd20567744bf155fb2a95c97717fa56", install=TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
# Write code here to load the ACS variables
this.year = 2019

vars <- load_variables(year = 2019,
                       dataset = "acs5",
                       cache = TRUE)

dim(vars)


# acs_2019 <- get_acs(geography = "county",
                    # variables = c(medincome = "B19013_001"),
                    # state = "MD",
                    # year = 2019)
```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
#Write code here to read the RDS file

ppp_loans_july <- read_rds("data/lab_08.rds")
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first.
**A1** The Maryland jurisdiction with the lowest median age is Baltimore city, Maryland. They have an estimate median age of 35.4.

```{r}
acs_2019_age <- get_acs(geography = "county",
                    variables = c(medage = "B01002_001"),
                    state = "MD",
                    year = 2019)
acs_2019_age %>% 
  arrange(estimate)

```

**Q2** Which Maryland counties have a median age of at least 50?
**A2** Talbot County and Worcester County

```{r}
acs_2019_age %>% 
  filter(estimate > "50")
```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? Start with the code you wrote for Q4 and add a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3** Prince George's County and Montgomery County has the highest percentage of 5-17 year-olds that speak Spanish at home. 

```{r}
acs_2019_5_17 <- get_acs(geography = "county",
                         variables = c(medkids = "B07001_003"),
                         state = "MD",
                         year = 2019)

acs_2019_spanish <- get_acs(geography = "county",
                    variables = c(medkidspan = "B06007_003"),
                    state = "MD",
                    year = 2019)

acs_2019_spanish %>%
  inner_join(acs_2019_5_17, by=c("GEOID", "NAME")) %>%
  rename(estimate_kid_span=estimate.x, estimate_kids=estimate.y) %>%
  select(-variable.x, -variable.y, -moe.x, -moe.y) %>%
  mutate(percentage = (estimate_kid_span/estimate_kids)*100) %>%
  arrange(desc(percentage))


```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4** Garrett County has the lowest percentage of 5-17 year olds that speak Spanish at home. One can possibly assume that the population of Spanish speaking kids (or even the population of Spanish people) is relatively low Garrett County. 

**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to create separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5** The 5-digit zip code that had the most removed PPP applications was 20850. The percentage of the Black population in that zip code is 13.34%.     

```{r}
# write.table(ppp_loans_july, file = "lab_08.csv", row.names = FALSE, sep = ",")

ppp_loans_july <- ppp_loans_july %>%
  mutate(zip5 = str_sub(zip, 1, 5)) 

top_zip <- maryland_ppp %>%
 group_by(zip5) %>%
 summarise(total=n()) %>%
 arrange(desc(total)) 

acs_2019_black <- get_acs(geography = "zcta",
                    variables = c(medblack = "B02001_003"),
                    state = "MD")

acs_2019_20850 <- get_acs(geography = "zcta",
                          variables = c(medzip = "B01001_001"),
                          state = "MD")

acs_2019_zipcodes <- acs_2019_20850 %>% 
  inner_join(acs_2019_black, by=c("GEOID", "NAME")) %>%
  rename(estimate_zip=estimate.x, estimate_zip_black=estimate.y) %>%
  select(-variable.x, -variable.y, -moe.x, -moe.y) %>%
  mutate(percentage = (estimate_zip_black/estimate_zip)*100)
  
acs_2019_black_zip <- top_zip %>% left_join(acs_2019_zipcodes, by=c('zip5'='GEOID')) %>% arrange(desc(total))

acs_2019_black_zip %>% 
  head(1)
```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed application in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6** First off, six of these zip codes are located in Montgomery County (20850, 20852, 20910, 20814, 20904, 20878). The zip codes with a Black population percentage of 47% or higher in only Baltimore (21117), Prince George's and one Montgomery County zip code (20904). But the two zip codes (20774 and 20706) with the highest black population are both located in Prince George's County. The other zip code outside of Montgomery County in this list is Anne Arundel County (21401).  

Some conclusions I would draw from this and bring to my editor would be that six of these zip codes (which are all not located in Montgomery County, saying this because that can be confusing) that have a Black population percentage of 30% or less had a lot of ppp loan applications removed. Two Prince George's County zip codes has the two highest Black population, and also had a lot of ppp loan applications removed. If I was told to find an angle and write a story on this, I would probably stick with the Prince George's county zip codes and wonder why such is the case. 

```{r}
acs_2019_black_zip %>% 
  arrange(desc(total)) %>% 
  head(10)
```

**Q7** What other columns from the PPP data could be useful in reporting on what happened to this removed loans?
**A7** Possibly the race column, because a lot of the fields were left unanswered... but specifically my question is what caused these applications to be removed? I don't know which column would give that hint, but I think that would be helpful when writing the story. 
