---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
allegations_nypd <- read_csv("data/allegations_202007271729.csv")

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: "Of at least 77 allegations made against him in 26 separate CCRB complaints, 29 were unsubstantiated; five were “unfounded,” meaning investigators concluded the incident never took place; and 27 were “exonerated,” meaning the conduct fell within patrol guidelines."

* **Analysis summary**: We first extracted Christopher Mccormack's information from the allegations_nypd dataframe by filtering his first and last name. Then we grouped by the CCRB complaint, which is the complaint_id column, and summarized the amount of allegations in each complaint. We found 72 allegations in 26 separate CCRB complaints. In order to find the amount of allegations that were unsubstantiated and exonerated, we filtered "Unsubstantiated" and "Exonerated" from the board_disposition column. We then grouped by the board_disposition and summarized the count of each, which showed us 27 exonerated allegations and 29 unsubstantiated allegations. 

As for the five allegations that were "unfounded," ProPublica explained that data excludes those five allegations. "We have published, and are releasing for download here, a version of the data that excludes any allegations that investigators concluded did not occur and were deemed unfounded." 

```{r}
# Put code to reverse engineer sentence here
mccormack_jackass <- allegations_nypd %>% 
  filter(first_name == "Christophe", last_name == "Mccormack")

mccormack_jackass %>% 
  group_by(complaint_id) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

mccormack_jackass %>% 
  filter(board_disposition == "Unsubstantiated"|board_disposition == "Exonerated") %>% 
  group_by(board_disposition) %>% 
  summarise(count = n())

# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: "In all, the CCRB substantiated 16 allegations lodged in six complaints, most involving the stops and searches of young men of color. No other high-ranking NYPD officer has amassed as many substantiated CCRB complaints."

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: “But many of the officers also said that it’s possible to be an effective cop without so many complaints. An overwhelming majority of the city’s 36,000 officers do not have a single substantiated complaint on their records; only 1 in 9 does.”

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: "Eighty-six of the roughly 420 officers in the department who currently hold a rank above captain — running precincts and other large commands and overseeing hundreds of officers — have tallied at least one misconduct allegation that was substantiated by the CCRB, meaning that investigators amassed enough evidence of offenses, ranging from bad language to pistol whippings, to say that they happened and broke patrol guidelines. The most common involved improper property entries and searches."

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: “McCormack has been accused of misconduct in at least 16 lawsuits, 11 of which settled for a combined $415,000. The allegations ranged from racial discrimination to strip-searches to false arrests.”

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-