---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
allegations_nypd <- read_csv("data/allegations_202007271729.csv")

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: "Of at least 77 allegations made against him in 26 separate CCRB complaints, 29 were unsubstantiated; five were “unfounded,” meaning investigators concluded the incident never took place; and 27 were “exonerated,” meaning the conduct fell within patrol guidelines."

* **Analysis summary**: We first extracted Christopher Mccormack's information from the allegations_nypd dataframe by filtering his first and last name. Then we grouped by the CCRB complaint, which is the complaint_id column, and summarized the amount of allegations in each complaint. We found 72 allegations in 26 separate CCRB complaints. In order to find the amount of allegations that were unsubstantiated and exonerated, we filtered "Unsubstantiated" and "Exonerated" from the board_disposition column. We then grouped by the board_disposition and summarized the count of each, which showed us 27 exonerated allegations and 29 unsubstantiated allegations. 

As for the five allegations that were "unfounded," ProPublica explained that data excludes those five allegations. "We have published, and are releasing for download here, a version of the data that excludes any allegations that investigators concluded did not occur and were deemed unfounded." 

```{r}
# Put code to reverse engineer sentence here
mccormack_jackass <- allegations_nypd %>% 
  filter(first_name == "Christophe", last_name == "Mccormack")

mccormack_jackass %>% 
  group_by(complaint_id) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

mccormack_jackass %>% 
  filter(board_disposition == "Unsubstantiated"|board_disposition == "Exonerated") %>% 
  group_by(board_disposition) %>% 
  summarise(count = n())

# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: "In all, the CCRB substantiated 16 allegations lodged in six complaints, most involving the stops and searches of young men of color." 

* **Analysis summary**: We filtered the allegations against MacCormack to isolate the substantiated allegations. This gave us a dataframe with 16 allegations. We grouped the allegations by Complaint ID and found six complaints.

We then filtered the substantiated allegations for "frisk" or "search" and found seven of the 16 were frisks or searches of black or Hispanic men, all but one in their 20s.

```{r}
# Put code to reverse engineer sentence here
mccormack_substantiated <- mccormack_jackass %>% 
  filter(str_detect(board_disposition, "^Substantiated"))

mccormack_substantiated %>% 
  group_by(complaint_id) %>% 
  summarise(total = n())

mccormack_substantiated %>% 
  filter(str_detect(allegation, "(Frisk|search)"))


# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: “But many of the officers also said that it’s possible to be an effective cop without so many complaints. An overwhelming majority of the city’s 36,000 officers do not have a single substantiated complaint on their records; only 1 in 9 does.” 

* **Analysis summary**: We filtered the data source for "substantiated" allegations and then grouped by "Unique ID" to get a count of officers with substantiated claims. We found 3,996 officers with substantiated claims, which is roughly 1 in 9 of 36,000 New York City police officers.

We could not verify the total count of 36,000 police officers from the database but double checked the figure with the NYPD website at [https://www1.nyc.gov/site/nypd/about/about-nypd/about-nypd-landing.page].

```{r}
# Put code to reverse engineer sentence here
all_cops <- allegations_nypd %>% 
  group_by(unique_mos_id) %>% 
  summarise(total = n())

substantiated_cops <- allegations_nypd %>% 
  filter(str_detect(board_disposition, "^Substantiated")) %>% 
  group_by(unique_mos_id) %>% 
  summarise(total = n())

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: "Eighty-six of the roughly 420 officers in the department who currently hold a rank above captain — running precincts and other large commands and overseeing hundreds of officers — have tallied at least one misconduct allegation that was substantiated by the CCRB, meaning that investigators amassed enough evidence of offenses, ranging from bad language to pistol whippings, to say that they happened and broke patrol guidelines."

* **Analysis summary**: 
We could not reproduce the figure that 86 of roughly 420 officers who hold a rank above captain had at least one substantiated misconduct allegation. After asking about the figure, we were told that the data used by ProPublica included cases that were closed pending litigation or taken over by internal disciplinary processes. These were not included in the publicly released data.

As a result, in the publicly released data set, we were able to identify 81 of the 86 officers with a rank above captain with at least one substantiated misconduct allegation.

```{r}
# Put code to reverse engineer sentence here
substantiated_allegations <- allegations_nypd %>% 
  filter(str_detect(board_disposition, "^Substantiated")) %>% 
  filter(rank_abbrev_now != "POM") %>% 
  filter(rank_abbrev_now != "POF") %>% 
  filter(rank_abbrev_now != "PO") %>% 
  filter(rank_abbrev_now != "SGT") %>% 
  filter(rank_abbrev_now != "SSA") %>% 
  filter(rank_abbrev_now != "SDS") %>% 
  filter(rank_abbrev_now != "LT") %>% 
  filter(rank_abbrev_now != "DT1") %>% 
  filter(rank_abbrev_now != "DT2") %>% 
  filter(rank_abbrev_now != "DT3") %>% 
  filter(rank_abbrev_now != "DTS") %>% 
  filter(rank_abbrev_now != "CPT") %>% 
  filter(rank_abbrev_now != "LCD") %>% 
  filter(rank_abbrev_now != "LSA")

substantiated_allegations %>% 
  group_by(unique_mos_id) %>% 
  summarise(total = n())
  
  

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: "The most common involved improper property entries and searches."

* **Analysis summary**: This is a sentence refers to "86" (81 in our findings) officers who hold a rank above captain. We used the substantiated_allegations dataframe and grouped by the allegations. We summarised the amount of allegations and arranged it in descending order. We found that the allegation, "premises entered and/or searched" had the most amount which was 18.  

```{r}
# Put code to reverse engineer sentence here

substantiated_allegations %>% 
  group_by(allegation) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total))
  

# Display results of code below this codeblock

```

-30-