---
title: "lab_11"
author: "sean mussenden"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab. You'll need tidyverse, tidycensus, janitor, sf and tigris, and you'll also need to install your census API key.

```{r}
library(tidycensus)
#census_api_key("dddd8ba3acd20567744bf155fb2a95c97717fa56", install=TRUE)
library(tidyverse)
library(sf)
library(janitor)
library(tigris)

```

## Load data

**Task** Create a codeblock and load the WV slice of the PPP loan data, which is in the data folder and is called lab_11.rds.

```{r}
wv_ppp_loans <- read_rds("data/lab_11.rds")
```

## Questions 

**Q1.** You are interested in doing some reporting on how West Virginia companies in the "Mining, Quarrying, and Oil and Gas Extraction" industry (https://www.naics.com/six-digit-naics/?code=21) used PPP loans during the pandemic (West Virginia is a big coal mining state).  

Then produce a choropleth map that shows West Virginia counties shaded according to the total amount of loans to that industry per 100,000 population. 

Use the magma color palette, and shade on a "log" scale. Also produce a table that shows the total amount of loans per 100,000 population for each county, ranked from highest to lowest, along with the total amount of loans and the total number of loans. 

Then answer these questions:

* Based on the map, what region of the state appears to have gotten more loans in this industry? What geographic pattern do you notice in the data?
* Based on the table, what county got the highest total loan amount? What was the total amount per 100K population, the total amount of loans (not population adjusted), how many loans were there, and what's the population?

**A1.** The North East region seems to have gotten more loans in this industry, specifically Kanawha County. Sounds weird, but going diagonally down in the middle of the state, from the straight tip that is at a right angle, they all have more loans than the left or right sides of the state. 

Harrison County got the highest loan amount. For Harrison county, the total amount of loans per 100k was 14,607,650.03, they received a total of 9,919,762.98, they had a total of 59 loans, and their population is 67,908.   

```{r}
wv_ppp_loans

mining_wv_ppp_loans <- wv_ppp_loans %>% 
  filter(str_detect(naics_code, "^21")) %>% 
  rename("county" = "project_county_name") %>% 
  group_by(county) %>% 
  summarise(total_loans = sum(amount)) %>% 
  mutate(county = str_squish(county))
  
wv_county_pop <- get_acs(geography = "county",
                         variables = c(population = "B01001_001"),
                         state = "WV",
                         geometry = TRUE)

wv_county_pop <- wv_county_pop %>% 
  rename("population" = "estimate") %>%
  rename("county" = "NAME") %>% 
  select(-GEOID, -variable, -moe) %>% 
  mutate(county = toupper(str_remove_all(county,", West Virginia|County"))) %>% 
  #it kept coming up as NA for the total_loans column... maybe because it's 47 counties with ppp loans associated with mining rather than all 55? 
  mutate(county = str_squish(county))
#used the code above to put all the counties in there and keep the NA for the ones without loans. 

# After noticing that there's only 47 counties with mining company loans, and 55 in the acs data, I did this below to make sure that the numbers are correct..? 
wv_ppp_loans %>% 
  filter(str_detect(naics_code, "^21")) %>%
  group_by(project_county_name) %>% 
  summarise(total = n()) %>%
  View()
#The numbers are correct.

mining_wv_ppp_loans_per_100k <- wv_county_pop %>% 
  left_join(mining_wv_ppp_loans, by = "county") %>% 
  mutate(loans_per_100k = (total_loans/population)*100000) %>% 
  mutate(total_loans = replace_na(total_loans,0)) %>% 
  mutate(loans_per_100k = replace_na(loans_per_100k,0))

ggplot() + 
  geom_sf(data=mining_wv_ppp_loans_per_100k, aes(fill=loans_per_100k)) + 
  theme_minimal() +
  scale_fill_viridis_b(option="magma", trans = "log") 


#The table.
mining_wv_ppp_loan_applications <- wv_ppp_loans %>% 
  filter(str_detect(naics_code, "^21")) %>% 
  rename("county" = "project_county_name") %>% 
  group_by(county) %>% 
  summarise(total_applications = n())

mining_wv_ppp_all <- mining_wv_ppp_loans %>% 
  left_join(mining_wv_ppp_loan_applications) 

mining_wv_ppp_all <- mining_wv_ppp_all %>% 
  left_join(wv_county_pop) %>% 
  mutate(loans_per_100k = (total_loans/population)*100000) %>% 
  arrange(desc(loans_per_100k))
#Hm. The NA's don't come up in this one...

```

**Q2.** You want to understand WHY the spatial pattern you saw above exists.  So you grab a shapefile with the latitude and longitude coordinates of every mine permit in West Virginia from the state Department of Environmental Protection. 

Here's the url: https://tagis.dep.wv.gov/arcgis/rest/services/WVDEP_enterprise/mining_reclamation/MapServer/0/query?where=1%3D1&outFields=*&geometry=&geometryType=esriGeometryEnvelope&inSR=4326&spatialRel=esriSpatialRelIntersects&outSR=4326&f=json. 

(It's in a slightly different format than the .shp file in the lab. It's another type of shapefile called "geojson".  It will work just the same, though, despite that difference. You can read it directly into R without manually downloading it, by putting the url in the st_read() function we learned about in this week's prelab.) 

This data includes both currently operating mines and old ones that have closed.  You want to examine only mines that have a permit expiration date after Sept. 1, 2021. 

Make a map with a point for each one of these mine permits and put a blank map with the outline of West Virginia counties (which you can get from Tigris) behind it. 

What spatial pattern do you see? How does it compare to the spatial pattern in the first map you made?

**A2.** The patterns are very similar to the first map. The points on the second map are the same areas with the a higher amount of loans. These areas also have the highest request of loans (orange shows a lot more in the first map).

```{r}
operating_mines <- st_read("https://tagis.dep.wv.gov/arcgis/rest/services/WVDEP_enterprise/mining_reclamation/MapServer/0/query?where=1%3D1&outFields=*&geometry=&geometryType=esriGeometryEnvelope&inSR=4326&spatialRel=esriSpatialRelIntersects&outSR=4326&f=json")

after_sep1_2021_mines <- operating_mines %>% 
  filter(expiredate >= "2021/09/01") 

after_sep1_2021_mines %>%
  ggplot() + 
  geom_sf() + 
  theme_minimal()

counties <- counties() 
glimpse(counties)

wv_counties <- counties %>% 
  filter(STATEFP == "54")

wv_counties %>%
  ggplot() + 
  geom_sf() + 
  theme_minimal()

ggplot() + 
  geom_sf(data=wv_counties, fill = "white") + 
  geom_sf(data=after_sep1_2021_mines) +
  theme_minimal()

```

**Q3.** To confirm the spatial comparison you identified in question 2, let's create a new map that layers the mine location points on top of our PPP choropleth map we created in question 1. What do you notice about the amount of mines in the county that has the highest total loan amount per 100K?

**A3.** There are no mines in that county - which is Ritchie County. Well, there are two touching the outsides of that county. That's crazy. Probably some fraudulent activity going on?

```{r}
ggplot() + 
  geom_sf(data=mining_wv_ppp_loans_per_100k, aes(fill=loans_per_100k)) + 
  geom_sf(data=after_sep1_2021_mines) +
  theme_minimal() + 
  scale_fill_viridis_b(option="magma", trans = "log") 

```

**Q4.** Starting from the original WV ppp dataframe, examine the loans to companies in "Mining, Quarrying, and Oil and Gas Extraction" in the county with the highest total loan amount per 100K (you can use the table you created in question 1 to identify the top county)?
* What single company accounts for the biggest chunk of the total loan amount? 
* What specific industry are they in, based on their 6-digit NAICS code? Look them up on the web. What services do they provide?
* How many jobs did they save, according to the data?
* What's your hypothesis for why a company in their specific industry would need to seek a PPP loan during the pandemic?

**A4.** Hall Drilling LLC accounts for the biggest chunk of the total loan amount. They are specifically in the Drilling Oil and Gas Wells industry. They provide trucking, drilling, production installation, construction and roustabout services. They saved 387 jobs. I would assume COVID was the main reason why because COVID stoped a lot of things. I just don't know how COVID stopped/affected that specific industry. No one was leaving their houses, so there wasn't anything to move, hence no need for any trucking services. I think that any building plans for whatever kind of institution was put on pause/halted, so construction services stopped. Then the roustabout services like pressure washing and fence and gate installation, was also put on a pause as well because COVID and fear of interacting with anyone. So, no money was coming in. 

```{r}
ritchie_wv_ppp_loans <- wv_ppp_loans %>% 
  filter(str_detect(naics_code, "^21")) %>% 
  filter(project_county_name == "RITCHIE") %>% 
  arrange(desc(amount))
```
-30-
