---
title: "pre_lab_05.Rmd"
author: "sean mussenden"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructor demonstration points to hit
1. Ensure students install refinr package and load it. 
2. Show students how to download and install Open Refine and confirm install before they leave class. 


## Chapter 11

### Task 1: Install refinr package
**Task** Run the following code in the gray-colored codeblock below to install refinr.

```{r}
install.packages('refinr')
```

### Task 2: Load libraries and settings
**Task** Run the following code in the gray-colored codeblock below to load the tidyverse, janitor, and refinr and  turn off scientific notation.  
```{r}
# turn off sci notation
options(scipen=999)
library(tidyverse)
library(lubridate)
library(refinr)
```

### Task 3: Load data
**Task** Run the following code to load the zipped-up Maryland PPP loans data. How many records are in this table?
**Answer** There are 195,865 records.

```{r}
# Load data
md_loans <- read_csv("data/ppp_loans_md.csv.zip")

# Display it
md_loans

```

### Task 4: Group and count to identify problems in city column
**Task** Run the following code to group and count records in the city column, and filter to examine only cities that start with A. Name a problem that you see that is preventing proper grouping.
**Answer** There are cities with the exact same name, but they're not grouped together because some one variation of a name can be all upper case and the other variation can be lower case or title case, or vice versa. For example: "Abell" and "ABELL." "Aberdeen," "ABERDEEN" and "aberdeen." 

```{r}
# Now let's try and group and count the number of loans by city. To make it a bit more managable, let's use another string function from `stringr` and filter for cities that start with the uppercase "A" or lowercase "a" using the function `str_detect()` with a regular expression.  

# The filter function in the codeblock below says: look in the city column, and pluck out any value that starts with (the "^" symbol means "starts with") a lowercase "a" OR (the vertical "|", called a pipe, means OR) an uppercase "A".  

md_loans %>%
  group_by(city) %>%
  summarise(
    count=n()
  ) %>%
  filter(str_detect(city, "^a|^A")) %>%
  arrange(city)
```

### Task 5: Use refinr to clean the city_clean field. 
**Task** Run the following code to make a copy of the city field called city_clean, then use the `key_collision_merge` function from refinr to attempt to standarize similar values. Store it as a new object called clean_md_loans.  Click the black arrow on the table to examine the city and city_clean columns.  Looking at the results on the first page only (first 10 rows), name two changes the refinr package made, by comparing city to city_clean.
**Answer** Two of the changes the refinr package made include: "PHOENIX" to "Phoenix" and "WESTMINSTER" to "Westminster." They changed the city name from all uppercase to title case.  

```{r}
cleaned_md_loans <- md_loans %>%
  mutate(city_clean=key_collision_merge(city)) %>%
  select(id:city, city_clean, everything())

cleaned_md_loans

```

### Task 6: Filter cleaned_md_loans to examine cities that start with A 
**Task** Run the following code to group by city_clean and city, then filter for cities that start with A. Name three changes the refinr function made.  
**Answer** Three of the changes the refinr function made include: changing "aberdeen" to "Aberdeen," "ABERDEEN" to "Aberdeen," and "ABERDEEN PROVING GROUND" to "Aberdeen Proving Ground." The function essentially fixed the casing of the letters so that it will be easier to group and the computer wouldn't see the cities as seperate. 

```{r}
cleaned_md_loans %>%
  group_by(city_clean, city) %>%
  summarise(
    count=n()
  ) %>%
  filter(str_detect(city, "^a|^A")) %>%
  arrange(city)
```
### Task 7: Use a different refinr method to clean the city_clean field. 
**Task** Run the following code to use the n_gram_merge() function to clean the data, then group by city_clean and city to examine cities that start with A. Look at the differences. Does the changes it made vary substantially from the previous use of key_collision_merge?
**Answer** I don't see any differences. It seems to have done the exact same thing, but just a different code to do it. 

```{r}
cleaned_md_loans <- md_loans %>%
  mutate(city_clean=n_gram_merge(city)) %>%
  select(id:city, city_clean, everything())


cleaned_md_loans %>%
  group_by(city_clean, city) %>%
  summarise(
    count=n()
  ) %>%
  filter(str_detect(city, "^a|^A")) %>%
  arrange(city)


```
### Task 8: Follow the directions in the textbook to use Open Refine
**Task** Follow the directions in the textbook to upload ppp_loans_md.csv.zip to Open Refine. Follow the instructions to clean as many cities as the Open Refine clustering tool can identify.  In class, I will come around to look at the Open Refine project you created to do this. 
